---
title: "workflow_basic"
output: html_document
date: "2024-10-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

## Tutorial

### Basic syntax:

```{r}
x <- 5 # assignment
v <- c(1, 2, 3) # vector
v_2 <- v * 2 # vector operation
```

Use the variable defined in the code block: v has a value `r v`

### Functions

```{r}
# sequence
my_lst <- seq(from=1, to=10, by=2)
my_str <- "I really do not like the <- thing"
```

### Visualization

The super package tidyverse.

```{r}
# install.packages("tidyverse")
library(tidyverse)
```

Import datasets (e.g.: penguins dataset)

```{r}
# install.packages("palmerpenguins")
library(palmerpenguins)
```

Import color palette for plotting

```{r}
# install.packages("ggthemes")
library(ggthemes)
```

#### dataframe
```{r}
penguins
glimpse(penguins)
```

#### plotting

```{r}
ggplot(
  data = penguins, # load the dataframe
  mapping = aes(x = flipper_length_mm, y = body_mass_g) # x/y axis mapping to variables in dataframe species
) +
  geom_point(mapping=aes(color=species)) + # specify the type of plotting (line? scatter? bar chart?) + colors
  geom_smooth(method = "lm")
  
```

Note the correct level to define colors to not screw up the linear model fitting. The following get the colro right but the fitting not so ok.
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) +
  geom_point() +
  geom_smooth(method = "lm")
```

The final plot
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```

## Exercise

1. The variable defined is 'my_variable' while the one inspected is 'my_varıable', with 'i' replaced by 'ı'.

2. 
```{r}
library(tidyverse)
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) + geom_point() + geom_smooth(method = "lm")
```
3. Keyboard shortcut quick references. (Option + Shift + K)
4. 
```{r}
my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()
my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
```
The bar chart is saved. Since the bar chart is assigned to 'my_bar_plot'.

